Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-09-09 prop_letter.tex
prop_letter.tex Keep Updating Andy's note...

For example, a well-defined  
QIVPM, assuming the three intervals $\imposs$, $\necess$, and
$\unknown=\left[0,1\right]$, is:
\begin{equation}
\bar{\mu}_{1}\left(P\right)=\begin{cases}
\imposs\,, & \text{if }P=\mathbb{0}\text{ ;}\\
\necess\,, & \text{if }P=\mathbb{1}\text{ ;}\\
\unknown\,, & \text{otherwise.}
\end{cases}\label{eq:no-information-IVPM}
\end{equation}
This measure represents the beliefs of an experimenter with no prior
knowledge about the particular quantum system in question; any quantum
state would be consistent with $\bar{\mu}_{1}$ on any projectors, and
$\coreBorn\left(\bar{\mu}_{1},\events\right)$ is just the set of all
states. The subspace
$\events_{0}=\left\{ \mathbb{0},\proj{0},\proj{1},\mathbb{1}\right\} $
is a subspace of commuting events. Indeed restricting $\bar{\mu}_{1}$
to $\events_{0}$, $\bar{\mu}_{1}|_{\events_{0}}$, gives a classical
IVPM~$\bar{\mu}_{2}$:
\begin{equation}
\begin{aligned}\bar{\mu}_{2}(\mathbb{0}) & =\imposs & \bar{\mu}_{2}(\proj{0}) & =\unknown\\
\bar{\mu}_{2}(\mathbb{1}) & =\necess & \bar{\mu}_{2}(\proj{1}) & =\unknown
\end{aligned}
\end{equation}
If we identify the states ~$\proj{0}$ and $\proj{1}$ with the head
and the tail of a coin, respectively, then $\bar{\mu}_{2}$ represents
the situation in which that an experimenter knows nothing about a coin
before she tosses it. Moreover, for each $p\in\left[0,1\right]$, the
state~$p\proj{0}+\left(1-p\right)\proj{1}$ represents a classical
coin with probability~$p$ to get heads, and
$p\proj{0}+\left(1-p\right)\proj{1}$ is consistent with
$\bar{\mu}_{2}$ on $\events_{0}$ because any classical coin should be
consistent with the situation for which an experimenter has no prior
knowledge.

@@

Our generalization of quantum probability measures to QIVPMs allows us
to strengthen the scope of one of the fundamental theorems of quantum
information: the Kochen-Specker
theorem~\cite{kochenspecker1967,Redhead1987-REDINA,peres1995quantum,Jaeger2007,Held2016}. Our
finite-precision extension of the theorem will suggest a resolution to
the debate initiated by Meyer and Mermin on the relevance of the
Kochen-Specker to experimental, and hence finite-precision, quantum
measurements~\citep{PhysRevLett.83.3751,Mermin1999,Kent1999,HavlicekKrennSummhammerSvozil2001,SimonBruknerZeilinger2001,Cabello2002,Larsson2002,Appleby2002,BarrettKent2004,Appleby_2005,Spekkens2005,GuehneKleinmannCabelloEtAl2010,MazurekPuseyKunjwalEtAl2016}.
Specifically, the original Kochen-Specker theorem is formulated under
the assumption that all observables of a quantum mechanical system
have \emph{definite values at all times}~\cite{Held2016}, i.e., they
have infinitely precise values at all times. Our interval-valued
probability framework will allow us to state, and prove, a stronger
version of the theorem that holds even if the observables have values
that are only definite up to some precision specified by a parameter
$\delta$. Put differently, our stronger statement and theorem
quantifies Mermin's intuition~\citep{Mermin1999}:
\begin{quote}
  \ldots although the outcomes deduced from such imperfect
  measurements will occasionally differ dramatically from those
  allowed in the ideal case, if the misalignment is very slight, the
  statistical distribution of outcomes will differ only slightly from
  the ideal case.
\end{quote}

@@

The first step in our formalization is to introduce a family of QIVPMs
parameterized by an uncertainty~$\delta$ which we call
$\delta$-deterministic QIVPMs.

@@

\noindent The definition puts no restrictions on the choice of
intervals, only on the assignment of intervals to events. When
$\delta=0$, every event must be assigned a probability in either
interval $\imposs$ or $\necess$, i.e., every event is completely
determined with certainty. As $\delta$ gets larger, the QIVPM allows
for more indeterminate behavior. For example, assuming a set of
intervals including $\imposs$, $\necess$, and
$\{ [0,1/4], [1/4,1/2], [1/2,3/4], [3/4,1] \}$, the probability
measure that maps every event to either $\imposs$ or $\necess$ would
be a 0-deterministic QIVPM, and the probability measure that maps
every event to one of $\imposs$, $\necess$, $[0,1/5]$, or $[4/5,1]$
would be a $\delta$-deterministic QIVPM for every
$1/5 \le \delta \le 1$.

The expectation value of an observable relative to a 0-deterministic
QIVPM is fully determinate and is equal to one of the eigenvalues of
that observable. To see this, note that given an orthonormal basis
$\Omega=\{ \ket{\psi_0},\ket{\psi_1},\ldots\}$, a 0-deterministic
QIVPM must map exactly one of the projectors $\proj{\psi_i}$ to
$\necess$ and all others to $\imposs$. This is because by
Eq.~(\ref{eq:QIVPM-constraints}b), we have
$\bar{\mu}(\Sigma_i{\proj{\psi_i}})=\necess$ and by
Eq.~\ref{eq:QuantumInterval-valuedProbability-Inclusion}, we must have
one of the $\bar{\mu}(\proj{\psi_i})=\necess$ and all others mapped to
$\imposs$. Given any state~$\rho$ that is consistent with this QIVPM
on all the projectors in $\Omega$, we have by Eq.~(\ref{eq:hbar}) that
$\mu_{\rho}^{\mathrm{B}}$ must also map exactly one of the projectors
in $\Omega$ to 1 and all others to 0. If an observable has a spectral
decomposition along $\Omega$ then, by
Eq.~(\ref{eq:quantum-expectation}), its expectation value relative to
$\mu_{\rho}^{\mathrm{B}}$ is the eigenvalue $\lambda$ whose projector
is mapped to 1. It therefore follows, by
Eq.~(\ref{eq:quantum-interval-expectation-core}), that the expectation
value relative to the 0-deterministic~${\bar{\mu}}$ is fully
determinate and lies in the interval $[\lambda,\lambda]$.

>>>>

For example, a well-defined  
QIVPM, assuming the three intervals $\imposs$, $\necess$, and
$\unknown=\left[0,1\right]$, is:
\begin{equation}
\bar{\mu}_{1}\left(P\right)=\begin{cases}
\imposs\,, & \text{if }P=\mathbb{0}\text{ ;}\\
\necess\,, & \text{if }P=\mathbb{1}\text{ ;}\\
\unknown\,, & \text{otherwise.}
\end{cases}\label{eq:no-information-IVPM}
\end{equation}
This measure represents the beliefs of an experimenter with no prior
knowledge about the particular quantum system in question; any quantum
state would be consistent with $\bar{\mu}_{1}$ on any projectors, and
$\coreBorn\left(\bar{\mu}_{1},\events\right)$ is just the set of all
states. The subspace
$\events_{0}=\left\{ \mathbb{0},\proj{0},\proj{1},\mathbb{1}\right\} $
is a subspace of commuting events. Indeed restricting $\bar{\mu}_{1}$
to $\events_{0}$, $\bar{\mu}_{1}|_{\events_{0}}$, gives a classical
IVPM~$\bar{\mu}_{2}$:
\begin{equation}
\begin{aligned}\bar{\mu}_{2}(\mathbb{0}) & =\imposs & \bar{\mu}_{2}(\proj{0}) & =\unknown\\
\bar{\mu}_{2}(\mathbb{1}) & =\necess & \bar{\mu}_{2}(\proj{1}) & =\unknown
\end{aligned}
\end{equation}
If we identify the states ~$\proj{0}$ and $\proj{1}$ with the head
and the tail of a coin, respectively, then $\bar{\mu}_{2}$ represents
the situation in which an experimenter knows nothing about a coin
before she tosses it. Moreover, for each $p\in\left[0,1\right]$, the
state~$p\proj{0}+\left(1-p\right)\proj{1}$ represents a classical
coin with probability~$p$ to get heads, and
$p\proj{0}+\left(1-p\right)\proj{1}$ is consistent with
$\bar{\mu}_{2}$ on $\events_{0}$ because any classical coin should be
consistent with the situation for which an experimenter has no prior
knowledge.

@@

Our generalization of quantum probability measures to QIVPMs allows us
to strengthen the scope of one of the fundamental theorems of quantum
information: the Kochen-Specker
theorem~\cite{kochenspecker1967,Redhead1987-REDINA,peres1995quantum,Jaeger2007,Held2016}. Our
finite-precision extension of the theorem will suggest a resolution to
the debate initiated by Meyer and Mermin on the relevance of the
Kochen-Specker to experimental, and hence finite-precision, quantum
measurements~\citep{PhysRevLett.83.3751,Mermin1999,Kent1999,HavlicekKrennSummhammerSvozil2001,SimonBruknerZeilinger2001,Cabello2002,Larsson2002,Appleby2002,BarrettKent2004,Appleby_2005,Spekkens2005,GuehneKleinmannCabelloEtAl2010,MazurekPuseyKunjwalEtAl2016}.
Specifically, the original Kochen-Specker theorem is formulated
using a model quantum mechanical system that
has \emph{definite values at all times}~\cite{Held2016}, i.e., they
have infinitely precise values at all times. Our interval-valued
probability framework will allow us to state, and prove, a stronger
version of the theorem that holds even if the observables have values
that are only definite up to some precision specified by a parameter
$\delta$. Our approach provides a quantitative realization of
Mermin's intuition~\citep{Mermin1999}:
\begin{quote}
  \ldots although the outcomes deduced from such imperfect
  measurements will occasionally differ dramatically from those
  allowed in the ideal case, if the misalignment is very slight, the
  statistical distribution of outcomes will differ only slightly from
  the ideal case.
\end{quote}

@@

The first step in our formalization is to introduce a family of QIVPMs
parameterized by an uncertainty~$\delta$, which we call
$\delta$-deterministic QIVPMs.

@@

\noindent The definition puts no restrictions on the choice of
intervals, only on the assignment of intervals to events. When
$\delta=0$, every event must be assigned a probability either in the
interval $\imposs$ or in $\necess$, i.e., every event is completely
determined with certainty. As $\delta$ gets larger, the QIVPM allows
for more indeterminate behavior. For example, assuming a set of
intervals including $\imposs$, $\necess$, and
$\{ [0,1/4], [1/4,1/2], [1/2,3/4], [3/4,1] \}$, the probability
measure that maps every event to either $\imposs$ or $\necess$ would
be a 0-deterministic QIVPM, and the probability measure that maps
every event to one of $\imposs$, $\necess$, $[0,1/5]$, or $[4/5,1]$
would be a $\delta$-deterministic QIVPM for every
$1/5 \le \delta \le 1$.

The expectation value of an observable relative to a 0-deterministic
QIVPM is fully determinate and is equal to one of the eigenvalues of
that observable. To see this, note that given an orthonormal basis
$\Omega=\{ \ket{\psi_0},\ket{\psi_1},\ldots\}$, a 0-deterministic
QIVPM must map exactly one of the projectors $\proj{\psi_i}$ to
$\necess$ and all others to $\imposs$. This is because by
Eq.~(\ref{eq:QIVPM-constraints}b), we have
$\bar{\mu}(\Sigma_i{\proj{\psi_i}})=\necess$ and by inductively applying
Eq.~\ref{eq:QuantumInterval-valuedProbability-Inclusion} , we must have
one of the $\bar{\mu}(\proj{\psi_i})=\necess$ and all others mapped to
$\imposs$. Given any state~$\rho$ that is consistent with this QIVPM
on all the projectors in $\Omega$, we have by Eq.~(\ref{eq:hbar}) that
$\mu_{\rho}^{\mathrm{B}}$ must also map exactly one of the projectors
in $\Omega$ to 1 and all others to 0. If an observable has a spectral
decomposition along $\Omega$ then, by
Eq.~(\ref{eq:quantum-expectation}), its expectation value relative to
$\mu_{\rho}^{\mathrm{B}}$ is the eigenvalue $\lambda$ whose projector
is mapped to 1. It therefore follows, by
Eq.~(\ref{eq:quantum-interval-expectation-core}), that the expectation
value relative to the 0-deterministic~${\bar{\mu}}$ is fully
determinate and lies in the interval $[\lambda,\lambda]$.
